---
title: "Analysing Google Scholar data from R"
author: "Andi Fugard"
date: "20 Dec 2020"
output: html_document
---

This page shows how to plot citations data using the `scholar` package by Guangchuang Yu et al. and the `tidyverse`.


Load the packages:


```{r message=FALSE, warning=FALSE}
library(scholar)
library(tidyverse)
```

I'll use my citations data:

```{r}
andi_id <- "xrY7bFYAAAAJ"
pubs <- get_publications(andi_id)
```

Now `pubs` has a bunch of info on each of my papers:

```{r}
names(pubs)
```

Here's the current year, which will come in handy later.

```{r}
current_year <- as.numeric(format(Sys.Date(), "%Y"))
```

It was `r current_year` when I ran this.

Now let's look at citations each paper got by year. Get all citation data:

```{r}
all_cites <- map_dfr(pubs$pubid,
                       ~ get_article_cite_history(andi_id, .x))
```

This has one row per year and original cited paper:

```{r}
head(all_cites)
```

Now glue the original paper info back on. First reduce it a bit:

```{r}
red_pubs <- pubs |>
  mutate(paper_year = year,
         total_cites = cites) |>
  select(title, author, paper_year, journal, pubid, total_cites)
```

Join:

```{r}
all_cites_pub <- all_cites |>
  left_join(red_pubs)
```

Next, we need to add in implicit zero-cites where a year is missing, from the publication year up to current year.

This function does it for one paper's cite info:

```{r}
complete_paper_cites <- function(one_paper) {
  expand_years <- setdiff(unique(one_paper$paper_year):current_year,
                          one_paper$year)
  complete(one_paper,
           year = expand_years,
           cites = 0,
           pubid,
           title,
           author,
           paper_year,
           journal,
           total_cites) |>
    arrange(year)  
} 
```

Give it a quick test using this input:

```{r}
test_input <- all_cites_pub |>
  filter(pubid == "43bX7VzcjpAC")

test_input |>
  select(paper_year, year, cites)
```

Here's the output:

```{r}
complete_paper_cites(test_input) |>
  select(paper_year, year, cites)
```

Looks good, so tidy up:

```{r}
rm(test_input)
```

Now do this for all papers:

```{r}
all_cites_filled <- all_cites_pub |>
  group_split(pubid) |>
  map_dfr(complete_paper_cites) |>
  mutate(age = year - paper_year)
```

Calculate the cumulative citation sums:

```{r}
all_cites_cum <- all_cites_filled |>
  group_by(pubid) |>
  mutate(cum_sum = cumsum(cites))
```

Plot for all papers...

```{r dpi = 600}
all_cites_cum |>
  ggplot(aes(x = year, y = cum_sum, colour = pubid)) +
  geom_line(linewidth = 1, alpha = 0.5) +
  theme(legend.position = "none") +
  labs(x = "Year",
       y = "Cumulative citations",
       title = "All the papers")
```

That's quite busy so try again.

```{r}
min_cites <- 50
```

```{r fig.height=4, fig.width=7, dpi = 600, warning=FALSE}
to_plot <- all_cites_cum |>
  filter(total_cites >= min_cites)

to_plot |>
  ggplot(aes(x = year,
             y = cum_sum,
             colour = reorder(str_trunc(title, 20), desc(total_cites)))) +
  stat_smooth(geom="line", linewidth = 1, alpha = 0.5) +
  theme(legend.position = "right",
        legend.directio = "vertical") +
  labs(x = "Year",
       y = "Cumulative citations",
       colour = "Paper",
       title = paste("Papers cited at least", min_cites, "times")) +
  scale_x_continuous(breaks = seq(min(to_plot$year),max(to_plot$year)+2, 2)) +
  scale_y_continuous(breaks = seq(0,ceiling(max(to_plot$cum_sum)/200)*200, 200))
```

```{r fig.height=4, fig.width=7, dpi = 600, warning=FALSE}
to_plot <- all_cites_cum |>
  filter(total_cites >= min_cites)

to_plot |>
  filter(age >= 0 & cum_sum > 0) |>
  ggplot(aes(x = age,
             y = cum_sum,
             colour = reorder(str_trunc(title, 20), desc(total_cites)))) +
  stat_smooth(geom="line", linewidth = 1, alpha = 0.5) +
  theme(legend.position = "right",
        legend.directio = "vertical") +
  labs(x = "Age (years)",
       y = "Cumulative citations",
       colour = "Paper",
       title = paste("Papers cited at least", min_cites, "times")) +
  scale_x_continuous(breaks = seq(min(to_plot$age),max(to_plot$age)+2, 1)) +
  scale_y_continuous(trans = "log2")
```

